---
title: "Regression"
description: "ML assignment blog 4"
date: 2023-12-02
categories: [Machine Learning] 
draft: false
---

In this blog, we'll use the GLM to perform simple linear regression. Below,you will find me using mtcars data for fitting a linear regression model between the variables weight and miles per gallon for car dataset.

## 1. Simple linear regression

-   **Definition:** Simple linear regression involves predicting a dependent variable (response) based on a single independent variable (predictor).

-   **Equation:** y=β0+β1x+ϵ, where y is the dependent variable, x is the independent variable, β0​ is the intercept, β1​ is the slope, and ϵ is the error term

```{r}

library(ggplot2)

# Load the mtcars dataset
data(mtcars)

# Explore the dataset
head(mtcars)
summary(mtcars)

# Scatter plot
plot(mtcars$wt, mtcars$mpg, main="Scatter Plot", xlab="Weight", ylab="Miles Per Gallon", col="blue")

# Fit the linear regression model
model <- lm(mpg ~ wt, data = mtcars)

# Display the summary of the model
summary(model)

# Visualize the regression line


ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  stat_smooth(method = lm)



```

## 2. Non-Linear Regression Models:

Linear regression models assume a linear relationship between the independent variables and the dependent variable. However, in many real-world scenarios, the relationship may not be strictly linear. Non-linear regression models are used when the relationship between variables is better described by a non-linear equation.

```{r}


# Load the mtcars dataset
data(mtcars)

# Fit a quadratic non-linear regression model
model <- lm(mpg ~ poly(hp, 2), data = mtcars)

# Generate predicted values
predictions <- predict(model, newdata = data.frame(hp = mtcars$hp))

# Plot the data and non-linear regression curve
plot(mtcars$hp, mtcars$mpg, main="Quadratic Non-Linear Regression", xlab="Horsepower", ylab="Miles Per Gallon", col="blue")

ggplot(mtcars, aes(x = hp, y = mpg)) +
  geom_point(color = "blue") +
  
  # Add a dashed line for the non-linear regression curve
  geom_line(aes(y = predictions), color = "red") +
   # Labels and title
  labs(title = "Quadratic Non-Linear Regression",
       x = "Horsepower",
       y = "Miles Per Gallon") 

# Display the model summary
summary(model)

```
