---
title: "Clustering"
description: "ML assignment blog 2"
date: 2023-12-02
categories: [Machine Learning] 
draft: false

---

## Install and call the required libraries


```{r, warning=FALSE}


library(tidyverse)
library(janitor)
library(palmerpenguins)
library(knitr)
library(skimr)

```
## Data summary

```{r, warning=FALSE}

data= penguins_raw %>% clean_names()

skim(penguins_raw)

```
```{r, warning=FALSE}

summary(data)


```
## Visualisation and correlation check

```{r, warning=FALSE}

library(GGally)

ggpairs(
  data = data,
  columns = c(9:14),
  diag = list(continuous = wrap("barDiag", color = "red")),
  upper = list(continuous = wrap("cor", bins = 40))
)



```
There is a high correlation observed between body_mass_g and flipper_length. So, I opt to remove the one of the two variables. I am removing the body_mass_g variable from the dataset.

## Data wrangling

```{r, warning=FALSE}

   penguins = data %>%
      mutate (
         id = row_number(),
         species = word (species, 1),
         culmen_length_mm = scale(culmen_length_mm),
         culmen_depth_mm = scale(culmen_depth_mm),
         flipper_length_mm = scale(flipper_length_mm)
         ) %>%
      select (id, species, island, sex, culmen_length_mm, culmen_depth_mm, flipper_length_mm) %>%
      drop_na (sex)

summary(penguins)


```
The NA's are dropped. ID is added to each row. The variables culmen_length_mm, culmen_depth_mm and flipper_length_mm are scaled. We will use PCA analysis to reduce the dimensionality of the dataset. We will only use culmen_length_mm, culmen_depth_mm and flipper_length_mm for the K-means clustering and 




